### 2017-03-08 paq/parenting_obs_paq.Rmd
added parent_ed analysis

---
title: "PAQ for parenting observational study"
author: "Emily"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Data preprocessing

Preliminaries.

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE)
suppressPackageStartupMessages(c("dplyr","langcog","tidyr","ggplot2","lme4"))
library(psych)
library(langcog)
library(tidyverse)
library(ggthemes)
library(rjson)
library(stringr)
library(forcats)
library(tibble)

select <- dplyr::select # masked by MASS

theme_set(theme_few())
```

Read in participant data.

```{r}
files <- dir("parenting_obs_paq_online/")
attitudes <- data.frame()

#from online version
for (f in files) {
  jf <- paste("parenting_obs_paq_online/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))

  attitudes_id <- data.frame(sid = f, 
                   sent = jd$data$sentence[2:25],
                   rating = jd$data$rating[1:24])

  attitudes <- bind_rows(attitudes, attitudes_id)}

attitudes$sid <- str_replace(attitudes$sid, "parenting_proj_", "")
attitudes$sid <- str_replace(attitudes$sid, ".json", "")

#from paper version
paq <- read.csv('parenting_obs_paq_paper.csv')
sents <- read.csv('sent_forms_paper.csv')

paq <- paq %>%
  gather(ques, rating, Q1:Q24)%>%
  mutate(sid = Sub_ID)%>%
  left_join(sents)%>%
  select(sid, sent, rating)

paq$rating <- as.integer(paq$rating)
attitudes$rating <- as.integer(attitudes$rating)

attitudes <- bind_rows(attitudes, paq)
```

Read in trial info and questionnaire labels.

```{r}
labels <- read.csv("sent_forms.csv")
labels$sent <- as.character(labels$sent)
```

Clean up labels.

```{r}
attitudes$sent <- str_replace_all(as.character(attitudes$sent), "[â‘”“’']", "")
```

Read in demographics.

```{r}
demo <- read.csv("../parenting_proj_emilyhembacher_demo2016.csv")

demo <- demo %>%
  mutate(sid = SID)%>%
  select(sid, ethnicity, parent_ed)

attitudes <- left_join(attitudes, demo)
```

# Data Frames and Exclusions

## Make data frames.

Questionnaire attitude means.

```{r}
dq <- attitudes %>%
  left_join(labels)

dq$rating <- as.numeric(dq$rating)

dq$rating[dq$reverse_code == 1] <- 6 - dq$rating[dq$reverse_code == 1]

atts <- dq %>%
  group_by(sid, category) %>% 
  summarise(rating = mean(rating))
```

# Questionnaire

Look at mean ratings across sentences.

```{r}
ms <- dq %>%
  group_by(category, short_sent, reverse_code) %>%
  multi_boot_standard(col = "rating", na.rm = TRUE) %>%
  arrange(category, desc(mean))

ms$short_sent_ord <- factor(ms$short_sent, 
                             levels = ms$short_sent)
```

Plot responses to individual questionnaire items.

```{r}
qplot(short_sent_ord, mean, col = category,
      ymin = ci_lower, ymax = ci_upper, pch = factor(reverse_code),
      geom = "pointrange",
      data = ms) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) +
  xlab("") + 
  ylab("Mean Rating") + 
  ylim(c(0,6)) + 
  scale_colour_solarized()
```

Plot mean subscale scores.

```{r}
atts_m <- dq %>%
  group_by(category) %>%
  multi_boot_standard(col = "rating", na.rm = TRUE) %>%
  arrange(category, desc(mean))

ggplot(atts_m, aes(x = category, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))
```
```{r}
dq$parent_ed <- as.factor(dq$parent_ed)

mss <- dq %>%
  group_by(sid, parent_ed) %>%
  filter(!is.na(parent_ed))%>%
  multi_boot_standard(col = "rating", na.rm = TRUE)

qplot(parent_ed, data=mss)

ms <- dq %>%
  group_by(parent_ed, category) %>%
  filter(!is.na(parent_ed))%>%
  multi_boot_standard(col = "rating", na.rm = TRUE)

ggplot(ms, aes(category, mean, fill=parent_ed)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) + 
  scale_fill_solarized()
```