### 2017-05-31 joint_attention/parenting_obs_JA.Rmd
updated analysis for JA

---
title: "Parenting Observational Study"
author: "Emily"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
---

<style type="text/css">
body, td {
   font-size: 14px;
}
code {
  font-size: 11px;
}
pre {
  font-size: 11px;
}
</style>

Data analysis of social referencing and uncertainty study.

# Data preprocessing

Preliminaries.

```{r echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
library(tidyverse)
library(lme4)
library(knitr)
library(markdown)
library(lmerTest)
library(stringr)
library(langcog)
theme_set(theme_bw())
```

Read in files and consolidate to the same directory.

```{r}
d.raw <- data.frame()
files <- dir("data/")

for (f in files) {
  jf <- paste("data/",f,sep="")
  jd <- read.csv(jf)

  SID <- str_replace(f, ".csv", "")
  line <- jd$JointAttention.ordinal
  phase_onset <- jd$JointAttention.onset
  phase_offset <- jd$JointAttention.offset
  pja <- jd$JointAttention.pJA
  cja <- jd$JointAttention.cJA
  bid <- jd$JointAttention.bids

  id <- data.frame(SID = SID,
                   line = line,
                   phase_onset = phase_onset,
                   phase_offset = phase_offset, 
                   pja = pja,
                   cja = cja,
                   bid = bid)

  l_pja <- id%>%
    mutate(phase_length = phase_offset - phase_onset)%>%
    filter(pja == 1)%>%
    group_by(SID, pja)%>%
    summarise(pja_length = sum(phase_length))%>%
    select(-pja)

  l_cja <- id%>%
    mutate(phase_length = phase_offset - phase_onset)%>%
    filter(cja == 1)%>%
    group_by(SID, cja)%>%
    summarise(cja_length = sum(phase_length))%>%
    select(-cja)

  c_pja <- id%>%
    group_by(SID)%>%
    summarise(pja = sum(pja))

  c_cja <- id%>%
    group_by(SID)%>%
    summarise(cja = sum(cja))

  c_bids <- id%>%
    group_by(SID)%>%
    summarise(bids = sum(bid))

  sd <- c_bids%>%
    left_join(c_cja)%>%
    left_join(c_pja)%>%
    left_join(c_cja)%>%
    left_join(l_pja)%>%
    left_join(l_cja)

sd$pja_length[is.na(sd$pja_length)] <- 0
sd$cja_length[is.na(sd$cja_length)] <- 0

sd <- sd %>%
  mutate(total_ja = pja + cja)%>%
  mutate(total_lja = pja_length + cja_length)

d.raw <- bind_rows(d.raw, sd)
}
```

Read in trial info and demographics.

```{r}
# demo <- read.csv("../parenting_obs_demo.csv")
conditions <- read.csv("conditions.csv")

d <- d.raw %>%
  left_join(conditions)%>%
  mutate(bids_tot = bids + cja + pja)
```

```{r}
#total  # episodes of JA
ms_total_ja <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "total_ja")

ggplot(ms_total_ja, aes(x = Condition, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) +
  xlab("Condition") + 
  ylab("Episodes Joint Attention") 
```

```{r}
#total time JA
ms_total_lja <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "total_lja")

#plot length of joint attention.
ggplot(ms_total_lja, aes(x = Condition, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) +
  xlab("Condition") + 
  ylab("Time Spent in Joint Attention") 
```

```{r}
#unsuccessful bids
ms_bids <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "bids")

ggplot(ms_bids, aes(x = Condition, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Unsuccessful bids for Joint Attention") 
```

```{r}
# total bids
ms_bids <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "bids_tot")

ggplot(ms_bids, aes(x = Condition, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total bids for Joint Attention") 
```

```{r}
#pja
ms_pja <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "pja")

ggplot(ms_pja, aes(x = Condition, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Episodes of Passive Joint Attention") 
```

```{r}
#cja
ms_cja <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "cja")

ggplot(ms_cja, aes(x = Condition, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Episodes of Coordinated Joint Attention") 
```

```{r}
#cja
ms_lcja <- d %>%
  filter(!is.na(Condition))%>%
  group_by(Condition) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = Condition, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total time Spent in Coordinated Joint Attention") 
```

```{r}
#cja
ms_lpja <- d %>%
  filter(!is.na(Condition))%>%
  group_by(Condition) %>%
  multi_boot_standard(col = "pja_length")

#plot 
ggplot(ms_lpja, aes(x = Condition, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total time Spent in Passive Joint Attention") 
```

#Analysis

T-tests
```{r}
t.test(d$bids[d$Condition == "exp"],d$bids[d$Condition == "con"])
t.test(d$bids_tot[d$Condition == "exp"],d$bids_tot[d$Condition == "con"])
t.test(d$cja[d$Condition == "exp"],d$cja[d$Condition == "con"])
t.test(d$pja[d$Condition == "exp"],d$pja[d$Condition == "con"])
t.test(d$pja_length[d$Condition == "exp"],d$pja_length[d$Condition == "con"])
t.test(d$cja_length[d$Condition == "exp"],d$cja_length[d$Condition == "con"])
t.test(d$total_ja[d$Condition == "exp"],d$total_ja[d$Condition == "con"])
t.test(d$total_lja[d$Condition == "exp"],d$total_lja[d$Condition == "con"])
```

Significant effect of condition on total number of bids and unsuccessful bids (most bids are unsuccessful).

Mixed effects models.

```{r}
bids <- lmer(bids_tot ~ Condition  +
                           (1 | SID) + (1 | Video),
                         data = d)
```

```{r eval = FALSE}
maximal_mod <- lmer(l_pja ~ condition * EL_score +
                           (condition * EL_score | SID), 
                         data = lmer_data)

maximal_mod <- lmer(l_cja ~ condition * EL_score +
                           (condition * EL_score | SID), 
                         data = lmer_data)

maximal_mod <- lmer(c_bids ~ condition * EL_score +
                           (condition * EL_score | SID), 
                         data = lmer_data)

maximal_mod <- lmer(c_pja ~ condition * EL_score +
                           (condition * EL_score | SID), 
                         data = lmer_data)

maximal_mod <- lmer(c_cja ~ condition * EL_score +
                           (condition * EL_score | SID), 
                         data = lmer_data)
```