### 2017-07-07 parenting_obs_lexdiv.Rmd
add analyses for prereg

---
title: "Parenting observation lexical diversity"
author: "Emily"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Data preprocessing

Preliminaries.

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE)
suppressPackageStartupMessages(c("dplyr","langcog","tidyr","ggplot2","lme4"))
library(psych)
library(langcog)
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)

select <- dplyr::select # masked by MASS
theme_set(theme_few())
```

Read in participant data.

```{r}
data <- read.csv("data.csv", header =TRUE)
dem <- read.csv("parenting_proj_emilyhembacher_demo2016.csv", header =TRUE)
conditions <- read.csv("joint_attention/conditions.csv")
load("paq/paq_demo.RData")

#fix ids
data$SID <- stringr::str_trim(data$SID)
data$SID <- stringr::str_replace(data$SID, "_cut.txt", "")
data$SID <- stringr::str_replace(data$SID, "8283_", "")
data$SID <- stringr::str_replace(data$SID, "283_", "")
data$SID <- stringr::str_replace(data$SID, "_cut_save.txt", "")
data$SID <- stringr::str_replace(data$SID, "_cut_save", "")
data$SID <- stringr::str_replace(data$SID, "_cutsave.txt", "")
data$SID <- stringr::str_replace(data$SID, "parenting_obs_", "0")
data$SID <- stringr::str_replace(data$SID, "_cut_m4a.txt", "")

data$SID <- stringr::str_replace(data$SID, "05116_05", "050116_05")
data$SID <- stringr::str_replace(data$SID, "05116_2", "050116_02")
```

Make data frames.

```{r}
d <- left_join(data, conditions)%>%
  left_join(dem)%>%
  transmute(sid=SID, 
            types = Type.count, 
            tokens = Token.count, 
            lexdiv = Lexical.diversity, 
            condition= Condition, 
            video = Video, 
            age = age,
            gender = gender,
            parent_ed = parent_ed)%>%
  left_join(ids)%>%
  filter(!is.na(condition))
```

Plots.

Lexical Diversity.
```{r}
ms_lex <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "lexdiv")

ggplot(ms_lex, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Lexical Diversity") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
```

Number of Tokens.
```{r}
ms_tok <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "tokens")

ggplot(ms_tok, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Tokens") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
```

Number of Word Types.
```{r}
ms_type <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "types")

ggplot(ms_tok, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Number of Word Types") +
  langcog::scale_colour_solarized()  +
  ggthemes::theme_few() 
```

Analyses.

Prepare data.
```{r}
lmer_data <- d %>%
  filter(!is.na(AA), !is.na(EL), !is.na(RR))%>%
  mutate(condition = factor(condition), 
         lexdiv = as.numeric(lexdiv),
         EL = as.numeric(langcog::scale(EL, scale=FALSE)),
         AA = as.numeric(langcog::scale(AA, scale=FALSE)),
         RR = as.numeric(langcog::scale(RR, scale=FALSE)),
         age = as.numeric(langcog::scale(age, scale=FALSE)),
         gender = as.factor(gender),
         video = as.factor(video))
```

Predicting lexical diversity based on experimental condition, PAQ, demographics. 
```{r}
maximal_mod <- lmer(lexdiv ~ condition *  EL  + condition * AA + condition * RR  + age + gender + parent_ed +
                           (1| video), 
                         data = lmer_data)
summary(maximal_mod)
```

Predicting the number of word tokens based on experimental condition, PAQ, demographics.  
```{r}
maximal_mod <- lmer(tokens ~ condition *  EL  + condition * AA + condition * RR  + age + gender + parent_ed +
                           (1| video), 
                         data = lmer_data)
summary(maximal_mod)
```

Predicting the number of word types based on experimental condition, PAQ, demographics.  
```{r}
maximal_mod <- lmer(types ~ condition *  EL  + condition * AA + condition * RR  + age + gender + parent_ed +
                           (1| video), 
                         data = lmer_data)
summary(maximal_mod)
```