### 2019-02-11 paper/cogsci2019/cogsci2019.tex
revised intro and added demo and reliability info

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission
\cogscifinalcopy

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Understanding the impact of electronically-delivered parenting advice}

\author{{\large\bf Hanwen Vivian~Zhang} (\texttt{vivian3@stanford.edu}) \\ {\large\bf George~Kachergis} (\texttt{george.kachergis@gmail.com}) \\ {\large\bf Michael C.~Frank} (\texttt{mcfrank@stanford.edu}) \\  Department of Psychology, Stanford University \\  Palo Alto, CA 94301 USA}

\begin{document}

\maketitle

\begin{abstract}
Early parenting practices play an important role in shaping the future
outcomes of young children (Hart \& Risley, 1995; Heckman, 2006). In
particular, high quality early interactions and early language input
appear to facilitate more effective language learning and higher levels
of school performance. The rise in electronic parenting applications
(``apps'') holds the promise of delivering low-cost, positive
interventions on parenting style to a variety of different populations.
Of special interest are the parents of very young children, who are
often difficult to reach in other ways. Yet little is known about the
effects of communicating to parents through app-based interventions.
MTLD, types and tokens, joint attention, proportion of
nouns/verbs/adjectives, affect, concreteness, mean length of
utterance\ldots{}

\textbf{Keywords:}
digital parenting advice; joint attention; lexical diversity; guided
play;
\end{abstract}

\section{Introduction}\label{introduction}

Children spend a large portion of their waking time at play in their
environment. Playing with objects may allow them to discover hidden
features, as well as to build a causal understanding of how the objects
interact. Beyond learning object features, the laws of physics, and
causal relationships (e.g., Schulz \& Bonawitz, 2007), during play
children practice motor skills that help them effectively navigate their
world (Singer, Golinkoff, \& Hirsh-Pasek, 2006). Social play can help
children learn about human relationships, both through imitation of
adult behaviors and by experiencing and learning to process emotional
events (Singer et al., 2006).

\subsection{Guided Play Scaffolds
Learning}\label{guided-play-scaffolds-learning}

Children's early play behaviors are often assisted by more skilled and
knowledgeable play partners such as their caregivers and older siblings
(Kaye, 1970). Under such expert guidance, children are encouraged and
motivated to engage in more advanced play, undertaking explorations that
push the boundaries of what they would be able to do unaided (Vygotsky,
1980). These tutorial interactions have been shown to be important
components of child development (Wood, Bruner, \& Ross, 1976). Thus,
with the knowledge of both play and tutorial interactions, guided play,
which consists of both active and enjoyable activities as well as close
guidance of adults (Hirsh-Pasek \& Golinkoff, 2008) has drawn
researchers' interest. A study on preschoolers by Weisberg showed that
guided play scaffolds the environment while still allowing children to
maintain a large degree of control, and it outperforms
direct-instruction approaches in encouraging a variety of positive
academic outcomes (Weisberg, Hirsh-Pasek, \& Golinkoff, 2013). Another
study by Massey specifically found that guided play could facilitate
children's vocabulary and comprehensive language development and
subsequent literacy skills (Massey, 2013).

\subsection{Improving Parenting Practices and Language
Use}\label{improving-parenting-practices-and-language-use}

Thus, including guided play in parenting practices early on is crucial
and will help with children's later performances in academic and social
situations. To intervene in parenting practices, Suskind and others did
a parent-directed home-visiting intervention experiment in 2015 for
children of low socioeconomic status. They showed that parents in
experimental group have greater knowledge of language development and
this effect sustained four months after the intervention. However, more
interactive outcomes, like adult word tokens, conversational turn
counts, and child vocalization counts, though increased significantly
during the experiment, did not sustain after the intervention (D. L.
Suskind et al., 2015). That changes did not sustain could be due to the
intervention itself, or could merely be that the methods of
home-visiting is not sustainable enough for parents to easily and
constantly get parenting advice. Thus, new methods of delivering
parenting advice should be considered.

\subsection{Effectiveness of Digital
Delivery}\label{effectiveness-of-digital-delivery}

With the development of technology, digitally-delivered methods could
address some logistic barriers to face-to-face delivery methods.
Breitenstein has reviewed 11 studies on them and showed good promise for
the efficacy of interventions using digitally-delivered methods
(Breitenstein, Gross, \& Christophersen, 2014). However, the parent and
child outcomes assessed in the review (e.g.infant positive behaviors,
satisfaction, emotional symptoms etc.) did not address the nature or
quality of parent-infant interactions at a detailed level, like if the
interventions lead to children paying more attention, or vocabulary
changes in parents' language usages. Although digitally-delivered
ac-tivities are designed to promote learning and cognitive devel-opment,
it is unclear how they might affect these dimensions of parent-child
interactions. Thus, we want to conduct this experiment to explore if and
how digital scaffolding of activities affect the social and linguistic
characteristics of parent-child ineractions. The quality of parent-child
interactions can be measured by both the social engagement of parents
(e.g., joint attention to objects in the environment)(Bigelow, MacLean,
\& Proctor, 2004) and the quality of language (e.g., vocabulary
diversity) (Malvern, Richards, Chipere, \& Durán, 2004).

\section{Experiment 1}\label{experiment-1}

In Experiment 1, we invited parents of 6- to 24-month-old infants
visiting the Children's Discovery Museum in San Jose to complete
activities from the Kinedu app. Parents were randomly assigned to the
video group or the control group; parents in the video group watched a
video from the Kinedu app (matched to their child's age), and then
performed the activity with their child using the props from the video.
Parents in the control group did not watch a Kinedu video, instead they
were given the same props and were told to play with their infants as
they would at home. We found that compared to the control group, parents
who watched a Kinedu video spoke more words overall, but had lower
lexical diversity. They also made more bids for joint attention with
their infants, although these bids did not result in more episodes of
joint attention compared to the control group. In summary, following
digitally-scaffolded activities may cause parents to engage with and
speak more to children overall, but speak more repetitively.

\subsection{Method}\label{method}

\begin{CodeChunk}
\begin{CodeOutput}
[1] 45
\end{CodeOutput}
\end{CodeChunk}

\subsubsection{Participants.}\label{participants.}

60 infants (F=42, M=18) aged 6-24 months (20 6-11.9 month-olds, 20
12-17.9 month-olds, and 20 18-24 month-olds) and their parents
participated in a museum in northern California. We included infants who
were exposed to English at least 50 percent of the time (n= 58) or who
were exposed less but whose participating parent reported that they
primarily speak English with their child at home (n = 2). Sixty-one\% of
participants (n=37) had been exposed to 2 or more languages as indicated
by their parent. Parents identified their children as White (n=25),
Asian (n=11), African American/Black (n=2), Biracial (n=12), other
(n=5), or declined to state (n=5). Fifteen parents reported their child
was of Hispanic origin. Parents tended to be highly educated, with a
highest level of education reported as high school (n = 5), some college
(n = 6), four-year college (n = 14), some graduate school (n = 1),
completed graduate school (n = 28) or declined to state (n=6).

\subsubsection{Materials.}\label{materials.}

Stimuli included videos from a commercial parenting application, Kinedu
Inc. The videos were designed to show activities parents could perform
with their child to foster cognitive and physical development, and were
targeted to the child's age and development. In each video, an adult and
child perform the activity while a narrator explains the activity and
its purpose. We selected two videos for each of three age groups in our
sample (6-11.9 months, 12-17.9 months, 18-23.94 months). More
information about the specific videos is available in the Appendix.
Participants were also given a set of props corresponding to those in
the video they watched, so that they could complete the activity. The
props associated with each video are listed in the Appendix.

Participants were randomly assigned to either the \emph{Video} condition
or the \emph{Control} condition. Participants in the Activity Video
condition were assigned to one of the two videos available for their
child's age group. The No-Video Control condition was yoked to the
Activity Video condition such that for every participant in the Activity
Video condition who saw a particular video and received the associated
props, a participant in the No-Video Control condition received the same
props but did not watch the corresponding video. Parents also completed
the Parenting Attitudes Questionnaire (PAQ). The PAQ measures parents'
attitudes about parenting and child development along three dimensions:
rules and respect, early learning, and affection and attachment.

\subsection{Procedure.}\label{procedure.}

After providing informed consent, parents in the Activity Video
condition watched the assigned video on a laptop with headphones. To
ensure that parents could give the video their full attention, the
experimenter played with the infant with a set of toys (different from
the experimental props used in the study) while the video was being
played. Immediately following the video, each parent-child dyad was
provided with the props to complete the activity they had viewed. The
toys were placed on a large foam mat, and parents were instructed to sit
on the mat with their child and re-create the activity they had viewed
for a period of three minutes. In the No-Video control condition, after
consenting parents were told to play with their child as they would at
home with the provided props for a period of three minutes. They were
not given any additional instructions about how to use the props.

In both conditions, two video cameras were used to record the play
session from different angles, and parents were fitted with a wireless
audio recorder to record their speech to their child. After three
minutes of play had elapsed, parents were told they could stop playing
and cameras and audio were turned off. Parents were then asked to
complete the PAQ before being debriefed.

\subsubsection{Joint Attention Coding
Procedure.}\label{joint-attention-coding-procedure.}

The video of each session was manually coded for episodes of joint
attention using the Datavyu software. The video taken at floor level was
coded by default, but the other video was referred to if the
participants were not visible or if there was a technological error with
the first camera. For each participant we coded for episodes of
coordinated joint attention, episodes of passive joint attention, and
parental bids for joint attention. We defined parental bids for joint
attention as any attempt to initiate joint attention (i.e label, point,
or otherwise draw attention to an object) that does not result in
passive or coordinated joint attention. If more than 3 seconds elapsed
between bids, they were coded as separate attempts. An episode of joint
attention was considered passive if both participants visually focused
on an object for a minimum of 3 seconds but the child did not
acknowledge the parent. If either participant looked away from the
object for less than 3 seconds and then returned to the same object it
was considered part of the same period of joint attention. Episodes
joint attention were considered coordinated if both participants
visually focused on an object for a minimum of 3 seconds and at some
point in the interaction the child indicated awareness of interaction
with some overt behavior toward the parent. This could be looks to the
parent's face, gestures, vocalizations, or turn-taking. If either
participant looked away for less than 3 seconds and then returned to the
same object, it was coded as part of the same period of joint attention.
A second coder independently coded a third of the videos (i.e., 20 of
the 60 videos, approximately equally distributed across ages) to
establish reliability. The two coders achieved reliability with ICC =
0.80 with 95\% confident interval = 0.57-0.92 (p\textless{}0.05) for
number of parent bids; ICC = 0.20 with 95\% confident interval =
-0.26-0.58 for number of passive joint attention established; ICC = 0.66
with 95\% confident interval = 0.32-0.85 (p\textless{}0.05) for number
of coordinated joint attention established; ICC = 0.24 with 95\%
confident interval = -0.21-0.61 for length of passive joint attention,
and ICC = 0.62 with 95\% confident interval = 0.27-0.83
(p\textless{}0.05) for length of passive joint attention.

\section{Results}\label{results}

The data was analyzed according to our preregistration:
\href{https://osf.io/2bpdf/}{link}.

\subsection{Lexical Diversity}\label{lexical-diversity}

We first calculated Type/Token Ratio (TTR) for each transcript. ..

Fig 1 - TTR, maybe also Total Tokens and Total Types (if subfigures fit)
(otherwise a table with mean and SD of tokens, types, and TTR) However,
note that TTR is a measure that is correlated with the length of a text.
\ldots{} MTLD is a better measure of lexical diversity (McCarthy \&
Jarvis, 2010).

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/e1lex_div-1}

}

\caption[Mean lexical diversity scores by condition (left]{Mean lexical diversity scores by condition (left: Type/Token ratio, right: MTLD) in Experiment 1.}\label{fig:e1lex_div}
\end{figure}
\end{CodeChunk}

Word tokens and word types

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/e1token_type-1}

}

\caption[Mean number of word types and word tokens by condition in Experiment 1]{Mean number of word types and word tokens by condition in Experiment 1.}\label{fig:e1token_type}
\end{figure}
\end{CodeChunk}

\subsection{Joint Attention}\label{joint-attention}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/e1ja-graphs-1}

}

\caption[Mean number of bids and episodes of Joint Attention in Experiment 1]{Mean number of bids and episodes of Joint Attention in Experiment 1.}\label{fig:e1ja-graphs}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/e1ja-graphs-pass-coord-1}

}

\caption[Average number of passive and coordinated episodes of JA in Experiment 1]{Average number of passive and coordinated episodes of JA in Experiment 1.}\label{fig:e1ja-graphs-pass-coord}
\end{figure}
\end{CodeChunk}

\subsection{Discussion}\label{discussion}

Both the number of tokens and types are higher in the experimental
condition, while lexical diversity (Type/Token ratio) is higher in the
control condition. Parents may be relatively more repetetive in the
experimental condition since they are attempting to stick to a specific
prescribed task, but they talk more overall. Demographics and PAQ do not
interact with condition, but there is a marginal effect of RR score on
lexical diversity (lower diversity for higher RR scores), and marginal
effects of parent education on word types and tokens (more types and
tokens for higher parent education).

There is a main effect of condition on total bids for joint attention.
Parents in the experimental condition (i.e., those who saw a video
demonstrating an activity) made a greater number of bids for joint
attention with their child. There was no effect of condition on the
number of episodes of either passive or coordinated joint attention, or
the duration of these episodes. There is a marginal effect of gender on
bids for joint attention, with parents of males producing more bids.
There is a marginal interaction between RR scores and condition on
passive joint attention, such that the experimental condition increased
the number of episodes of PJA to a greater extent for people with high
RR scores. While the electronically-delivered parenting advice increased
the number of bids for joint attention by parents, it did not
significantly effect the number or duration of episodes of joint
attention. One possibility is that child variables had a comparatively
larger impact on the attainment of joint attention.

\section{Experiment 2}\label{experiment-2}

In Experiment 2 we attempt to replicate the findings from Experiment 1
with a restricted number of preregistered predictions. We will
additionally include a second control condition, in which the same
activities are described in written form, rather than being demonstrated
in video. This manipulation will help determine what the contribution of
the video demonstration is in producing the observed effects on
parent-child interactions.

\subsection{Method}\label{method-1}

\subsubsection{Participants.}\label{participants.-1}

84 infants aged 12-24 months and their parents participated in the same
museum as Experiment 1. We included infants who were exposed to English
at least 75 percent of the time or who were exposed less but whose
participating parent reported that they primarily speak English with
their child at home. {[}specific demo info{]}

\subsubsection{Materials.}\label{materials.-1}

The design of Experiment 2 was similar to that of Experiment 1, except
that instead of a No-Video control condition, parents instead watched a
video that was generally related to child development research, but did
not give any specific instructions about how to interact with infants or
children. This was to control for the possibility that differences in
language output and joint attention in Experiment 1 could be due to
simply cuing parents to think about infants' learning and cognitive
development. The videos presented in the Control Video condition were
media clips (available on YouTube) of developmental psychologists
explaining their research interleaved with footage of infants or
toddlers engaged in developmental research studies. Thus, the content of
the videos superficially matched those in the Activity Video condition,
but did not suggest any particular activities. The videos were trimmed
to approximately match the average video length in the Activity Video
condition around 1.5 minutes.

\subsubsection{Procedure.}\label{procedure.-1}

The procedure for Experiment 2 matched that of Experiment 1, except that
parents in the Control Video condition watched a control video before
the play session. Consistent with the No-Video control condition in
Experiment 1, parents in the Control Video condition were told to play
with their child as they would at home, and were not given additional
instructions.

\section{Results}\label{results-1}

\subsection{Lexical Diversity}\label{lexical-diversity-1}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/e2lexdiv-1}

}

\caption[Mean lexical diversity scores by condition (left]{Mean lexical diversity scores by condition (left: Type/Token ratio, right: MTLD) in Experiment 2.}\label{fig:e2lexdiv}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/e2token-type-1}

}

\caption[Mean number of word types and word tokens by condition in Experiment 2]{Mean number of word types and word tokens by condition in Experiment 2.}\label{fig:e2token-type}
\end{figure}
\end{CodeChunk}

\subsection{Joint Attention}\label{joint-attention-1}

\section{Discussion}\label{discussion-1}

\subsection{Footnotes}\label{footnotes}

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. You can also use markdown formatting to include
footnotes using this syntax.\footnote{Sample of a markdown footnote.}

\subsection{Two-column images}\label{two-column-images}

You might want to display a wide figure across both columns. To do this,
you change the \texttt{fig.env} chunk option to \texttt{figure*}. To
align the image in the center of the page, set \texttt{fig.align} option
to \texttt{center}. To format the width of your caption text, you set
the \texttt{num.cols.cap} option to \texttt{2}.

\section{References}\label{references}

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

\hypertarget{refs}{}
\hypertarget{ref-Bigelow2004}{}
Bigelow, A. E., MacLean, K., \& Proctor, J. (2004). The role of joint
attention in the development of infants' play with objects.
\emph{Developmental Science}, \emph{7}, 518--526.

\hypertarget{ref-Breitenstein2014}{}
Breitenstein, S. M., Gross, D., \& Christophersen, R. (2014). Digital
delivery methods of parenting training interventions: A systematic
review. \emph{Worldviews on Evidence-Based Nursing}, \emph{11},
168--176.

\hypertarget{ref-Hart1995}{}
Hart, B., \& Risley, T. R. (1995). \emph{Meaningful differences in the
everyday experience of young american children}. Baltimore, MD: Brookes.

\hypertarget{ref-Heckman2006}{}
Heckman, J. J. (2006). Skill formation and the economics of investing in
disadvantaged children. \emph{Science}, \emph{312}(5782), 1900--1902.
\url{http://doi.org/10.1126/science.1128898}

\hypertarget{ref-Hirsh2008}{}
Hirsh-Pasek, K., \& Golinkoff, R. M. (2008). Why play= learning.
\emph{Encyclopedia on Early Childhood Development}, 1--7.

\hypertarget{ref-Kaye1970}{}
Kaye, K. (1970). Mother-child instructional interaction.
\emph{Unpublished Doctoral Thesis, Department of Psychology, Harvard
University}.

\hypertarget{ref-Malvern2004}{}
Malvern, D., Richards, B. J., Chipere, N., \& Durán, P. (2004). Lexical
diversity and language development.

\hypertarget{ref-Massey2013}{}
Massey, S. L. (2013). From the reading rug to the play center: Enhancing
vocabulary and comprehensive language skills by connecting storybook
reading and guided play. \emph{Early Childhood Education Journal},
\emph{41}, 125--131.

\hypertarget{ref-McCarthy2010}{}
McCarthy, P. M., \& Jarvis, S. (2010). MTLD, vocd-d, and hd-d: A
validation study of sophisticated approaches to lexical diversity
assessment. \emph{Behavior Research Methods}, \emph{42}(2), 381--392.

\hypertarget{ref-Schulz2007}{}
Schulz, L., \& Bonawitz, E. (2007). Serious fun: Preschoolers engage in
more exploratory play when evidence is confounded. \emph{Developmental
Psychology}, \emph{43}(4), 1045--1050.

\hypertarget{ref-Singer2006}{}
Singer, D. G., Golinkoff, R. M., \& Hirsh-Pasek, K. (Eds.). (2006).
\emph{Play = learning: How play motivates and enhances children's
cognitive and social-emotional growth}. New York, NY: Oxford University
Press.

\hypertarget{ref-Suskind2015}{}
Suskind, D. L., Leffel, K. R., Graf, E., Hernandez, M. W., Gunderson, E.
A., Sapolich, S. G., \ldots{} Levine, S. C. (2015). A parent-directed
language intervention for children of low socioeconomic status: A
randomized controlled pilot study. \emph{Journal of Child Language}.
\url{http://doi.org/10.1017/S0305000915000033}

\hypertarget{ref-Vygotsky1980}{}
Vygotsky, L. S. (1980). Mind in society: The development of higher
psychological processes.

\hypertarget{ref-Weisberg2013}{}
Weisberg, D. S., Hirsh-Pasek, K., \& Golinkoff, R. M. (2013). Guided
play: Where curricular goals meet a playful pedagogy. \emph{Mind, Brain,
and Education}, \emph{7}, 104--112.

\hypertarget{ref-Wood1976}{}
Wood, D., Bruner, J. S., \& Ross, G. (1976). The role of tutoring in
problem solving. \emph{Journal of Child Psychology and Psychiatry},
\emph{17}, 89--100.

\bibliographystyle{apacite}

\end{document}